Understanding the Causal Impact of Bike Lane Installation on Urban Cycling Adoption

### 2024-12-02 paper/paper.qmd
final paper edits

---
title: "Beyond Infrastructure: Understanding the Causal Impact of Bike Lane Installation on Urban Cycling Adoption"
subtitle: "A Difference-in-Differences Analysis of Toronto's Cycling Network, 2017-2023"
author: 
  - Steven Li
thanks: "Code and data are available at: [https://github.com/stevenli-uoft/Toronto_BikeShare_Causality](https://github.com/stevenli-uoft/Toronto_BikeShare_Causality)."
date: today
date-format: long
abstract: "We examine the causal effect of cycling infrastructure improvements on Bike Share Toronto ridership by analyzing how different types of bikeways influence system usage between 2017-2023. Using a difference-in-differences design with over 7.4 million rides across 1,191 bikeways, we find that infrastructure improvements led to significant increases in ridership, with post-implementation effects reaching 133.1 additional monthly rides per bikeway, translating to approximately 300,000 additional annual rides across all treated segments. Protected lanes demonstrate the strongest impact, generating 381.9 additional monthly rides - more than double the effect of on-road lanes (169.2 rides) and triple that of shared roadways (111.7 rides), while upgrades to existing infrastructure showed more immediate benefits than new installations. Our findings suggest that strategic implementation of protected infrastructure can substantially increase cycling activity, with implications for Toronto's planned addition of 100 kilometers of cycling infrastructure between 2025-2027. These results provide evidence-based guidance for cities working to encourage sustainable transportation through cycling infrastructure development."
format:
  pdf:
    toc: true
include-before: |
  \newpage
number-sections: true
bibliography: references.bib
---

@@

\newpage
# Introduction
Cycling infrastructure plays a crucial role in urban transportation, yet quantifying its impact on cycling adoption remains challenging. While protected bike lanes can reduce cycling injuries by up to 89% [@harris2020lane] and increase ridership by up to 171% on specific routes [@nacto2016], less is known about how different types of infrastructure improvements affect bike share system utilization. This knowledge gap is particularly relevant for Toronto, where rapid expansion of cycling infrastructure has coincided with dramatic growth in the bike share network from 2,854 to over 6,850 stations between 2016 and 2020 [@toronto_bike_share_expansion]. Understanding this relationship is critical as cities increasingly rely on bike share systems to complement public transit and provide sustainable transportation options.

We estimate the causal effect of cycling infrastructure changes on Bike Share Toronto ridership between 2019-2021, examining how protected lanes, on-road lanes, and shared roadways influence system usage. Using a difference-in-differences design, we analyze both new installations and infrastructure upgrades while controlling for broader trends in cycling adoption.

Our analysis reveals substantial increases in bike share ridership following infrastructure improvements. Post-implementation effects reach 133.1 additional monthly rides per bikeway (p < 0.01), translating to approximately 300,000 additional annual rides across all 186 treated segments - particularly notable as these improvements represent only 16% of the 1,191 bikeways in our study. Protected lanes demonstrate the strongest impact, generating 381.9 additional monthly rides (p < 0.001) - more than double the effect of on-road lanes and triple that of shared roadways. These gains occurred despite treated bikeways having higher initial ridership at 209.03 monthly rides, compared to control bikeways 85.59 monthly rides pre-treatment. This suggests infrastructure improvements can substantially increase cycling activity even on popular routes.

These findings carry important implications for urban transportation policy. Toronto's planned addition of 100 kilometers of cycling infrastructure between 2025-2027 [@cyclingNetworkPlan] presents an opportunity to enhance cycling adoption through strategic implementation. Our results challenge conventional assumptions about prioritizing underserved areas, as improvements to high-traffic routes generated significant additional ridership, highlighting the complex trade-offs in infrastructure resource allocation.

@@

# Data {#sec-data}
## Overview
This study draws on three datasets from the City of Toronto's Open Data Portal [@torontoOpenData]: the Cycling Network (Bikeways) data [@bikeWayData], and both Bike Share Ridership and Stations data [@bikeShareData]. Together, these datasets provide a comprehensive view of Toronto's cycling landscape, capturing infrastructure development from 2001-2023 and ridership patterns from 2017-2023. All data cleaning and analysis was conducted using R [@RCoreTeam2023], with key packages including `tidyverse` [@tidyverse], `dplyr` [@dplyr], and `sf` [@sf] for spatial analysis.

@@

A crucial methodological choice was the use of relative timing rather than calendar years. Due to bikeways having different upgrade or installation years, we align their treatment event year using a relative period. For treated bikeways, this creates a 60-month window: 24 months pre-treatment, 12 months during the treatment year, and 24 months post-treatment. Control bikeways receive randomly assigned pseudo-treatment years (2019, 2020, or 2021) to enable parallel analysis. This approach allows us facilitate direct comparison between treatment and control groups while accounting for the temporal distribution of infrastructure changes.

@@

## Outcome Variables
Our primary outcome variable is seasonally-adjusted monthly ridership for each bikeway (`monthly_rides_adj`). @fig-avg-monthly-ridership reveals several important patterns in ridership behavior. The raw ridership (black line) displays pronounced seasonal fluctuations, with consistent peaks during summer months and troughs in winter, reflecting Toronto's strong cycling seasonality. After applying seasonal adjustments (blue line), these cyclical patterns are smoothed while preserving the underlying growth trends. Both treatment and control groups show a general upward trend in ridership over the observation period, suggesting broader growth in cycling activity across Toronto's network.

@@

@fig-infra-type reveals distinct infrastructure patterns between groups. While protected lanes dominate both treatment and control groups, comprising approximately 50% of bikeways in each, the distribution of other types differs notably. The treatment group shows a higher proportion of on-road lanes relative to shared roadways, suggesting a deliberate strategy of upgrading less protected infrastructure. This aligns with Toronto's broader policy goal of enhancing cyclist safety through infrastructure improvements [@toronto_council_agenda].

@@

@@

Our analysis is ran in R [@RCoreTeam2023] using the `fixest` package [@fixest], which provides estimation methods for fixed effects models. This package is particularly suitable for our analysis as it handles multiple fixed effects efficiently and provides robust standard errors clustered at the bikeway level.

@@

### Parallel Trends Test
A crucial assumption of the DiD design is that treated and control bikeways would have followed parallel trends in the absence of treatment [@cuborican2024econometrics]. While this counterfactual cannot be directly tested, we can examine pre-treatment trends to assess the plausibility of this assumption. We implement this test by estimating:

@@

### Heterogeneity Analysis
We extend our analysis to examine treatment effect heterogeneity across different types of infrastructure changes. This investigation is crucial for understanding which interventions are most effective at increasing ridership. We estimate two separate models:

@@

These heterogeneity analyses reveal important nuances in the effectiveness of different infrastructure strategies. The particularly strong effect of protected lanes suggests that safety and separation from vehicle traffic may be key factors in encouraging Bike Share usage. Additionally, the stronger effects of upgrades compared to new installations may indicate the importance of improving infrastructure in areas where cycling activity already exists.

# Discussion {#sec-discussion}
Our difference-in-differences analysis of Toronto's cycling infrastructure changes between 2019-2021 reveals nuanced insights about how different types of cycling infrastructure improvements influence Bike Share utilization. The results demonstrate that while infrastructure investments broadly increase cycling adoption, the magnitude and timing of these effects vary substantially by implementation approach and infrastructure type.

@@

The significant increase in the post-treatment period – to 133.1 additional rides per month (p < 0.01) – reveals the true potential of infrastructure improvements once fully integrated into the transportation network. This impact from pre-treatment to post-treatment periods aligns with research from the University of Colorado Denver showing that "building safe facilities for cyclists is one of the biggest factors in road safety for everyone" [@MARSHALL2019100539]. The magnitude of this change suggests that infrastructure improvements don't just make existing cyclists' trips safer, they also alter the decisions behind mode of transportation for many residents.

@@

### Heterogenous Effect Findings
Our heterogeneous effects analysis reveals crucial insights about which interventions most effectively promote cycling. Protected lanes show remarkably strong effects, generating 381.9 additional monthly rides post-treatment (p < 0.001) – more than twice the impact of on-road lanes (169.2 rides, p < 0.001) and more than triple the effect of shared roadways (111.7 rides, p < 0.001). This substantial difference in effectiveness likely stems from:

@@

The comparison between upgrades and new installations provides equally important insights for implementation strategy. Upgrades to existing infrastructure generate more monthly ridership (152.7 adjusted rides), compared to newly-installing bike lanes (115.5 adjusted rides) post-treatment. Additionally, during the treatment period, upgrades to existing infrastructure show immediate effect to ridership (43.8 rides, p < 0.10), whereas new installations show no significant treatment-year effect. This pattern suggests that improving existing cycling routes may be more effective than expanding into new areas, particularly for short-term ridership gains. This finding aligns with Toronto's strategic focus during the same period as our study, on upgrading existing infrastructure to "higher-order" facilities, as outlined in the 2019-2021 Cycling Network Plan Update [@toronto_council_agenda].

@@

- **User surveys**: Qualitative research could help explain the mechanisms behind the stronger effects of protected lanes and how infrastructure influences mode choice decisions. Such research could particularly focus on understanding the "interested but concerned" demographic identified by @nacto2016 as crucial for expanding cycling adoption.
- **Network effects**: Future studies could examine how the benefits of infrastructure improvements spill over onto connecting routes and affect system-wide ridership patterns. This could help inform Toronto's goal of creating a more connected cycling network, as outlined in the city's 2025-2027 implementation program [@cyclingNetworkPlan].
- **Neighborhood-level analysis**: Future research could examine how infrastructure effects vary across Toronto's diverse neighborhoods. Such analysis would be particularly valuable as Toronto expands Bike Share into new areas, helping identify where different types of infrastructure improvements might have the greatest impact. This geographic heterogeneity analysis could also inform equity considerations in infrastructure deployment, ensuring benefits are distributed across the city's diverse communities.
- **Integration with public transit**: Studies could investigate how cycling infrastructure near transit hubs affects multi-modal trip patterns, particularly relevant given Toronto's efforts to integrate Bike Share with the TTC system [@ttcIntegration].

This research contributes to our understanding of how cycling infrastructure influences urban mobility patterns. The clear relationship between infrastructure quality and ridership, particularly the strong effect of protected lanes, provides evidence-based guidance for cities working to encourage sustainable transportation. As Toronto continues expanding its cycling network, these insights can help inform investment priorities and implementation strategies to maximize the impact of infrastructure improvements.

@@

Having this idealized survey would help us directly answer our research question by creating controlled variation in our sampling strategy, comparing treatment areas with new or upgraded bike lanes against control areas without changes. This design allows us to measure both first-stage effects of infrastructure changes on perceptions and intended behavior, and second-stage effects on actual ridership through follow-up surveys, while controlling for demographic and geographic factors. Through carefully designed questions about safety perceptions, route preferences, and time-value tradeoffs, we can identify the specific mechanisms through which infrastructure improvements affect ridership patterns. Additionally, our sampling of different user types enables analysis of heterogeneous effects across current, former, and potential users in various neighborhoods, providing a comprehensive understanding of how infrastructure changes influence Bike Share adoption.

@@

"Thank you for your time and participation..."Thank you for completing this survey! Your insights will help us better understand how bike infrastructure influences transportation choices in Toronto.

@@

As recommended by Stantcheva [-@Stantcheva2023] for testing persistence of effects, we will implement a comprehensive three-month follow-up survey. The key objective is to assess whether the stated preferences and intentions from the initial survey translate into actual behavioral changes, while also measuring any infrastructure-induced effects on ridership patterns. To maximize response rates, we will offer an additional $10 gift card incentive and keep the follow-up survey brief (under 3 minutes).

@@

## Tradeoffs and Limitations
Our survey methodology faces several important limitations that could affect the validity of our results. First, self-selection and non-response bias may skew our sample toward respondents who are already interested in cycling or have strong opinions about infrastructure, while our incentive structure might attract particular demographic groups, potentially missing key segments of the population. Our reliance on hypothetical scenarios and stated preferences introduces additional uncertainty, as actual behavior may differ significantly from reported intentions, especially regarding route selection and infrastructure usage. Despite careful design to minimize social desirability bias, respondents may still feel pressure to express support for sustainable transportation initiatives or understate safety concerns. Furthermore, temporal limitations of our three-month follow-up period may inadequately capture seasonal variations in cycling behavior and long-term adaptation to infrastructure changes, while the results may not generalize well to other urban contexts or represent the full diversity of potential users. These limitations suggest that while our survey can provide valuable insights into the relationship between infrastructure and ridership, findings should be interpreted alongside other data sources and methodologies.

>>>>

---
title: "Understanding the Causal Impact of Bike Lane Installation on Urban Cycling Adoption"
subtitle: "A Difference-in-Differences Analysis Shows Protected Bike Lanes Triple Bike Share Ridership in Toronto, 2017-2023"
author: 
  - Steven Li
thanks: "Code and data are available at: [https://github.com/stevenli-uoft/Toronto_BikeShare_Causality](https://github.com/stevenli-uoft/Toronto_BikeShare_Causality)."
date: today
date-format: long
abstract: "We examine the causal effect of cycling infrastructure improvements on Bike Share Toronto ridership by analyzing how different types of bikeways influence system usage between 2017-2023. Using a difference-in-differences design with over 7.4 million rides across 1,191 bikeways, we find that infrastructure improvements led to significant increases in ridership, with post-implementation effects reaching 133.1 additional monthly rides per bikeway, translating to approximately 300,000 additional annual rides across all treated segments. Protected lanes demonstrate the strongest impact, generating 381.9 additional monthly rides - more than double the effect of on-road lanes (169.2 rides) and triple that of shared roadways (111.7 rides), while upgrades to existing infrastructure showed more immediate benefits than new installations. Our findings suggest that strategic implementation of protected infrastructure can substantially increase cycling activity, with implications for Toronto's planned addition of 100 kilometers of cycling infrastructure between 2025-2027. These results provide evidence-based guidance for cities working to encourage sustainable transportation through cycling infrastructure development."
format:
  pdf:
    toc: true
include-before: |
  \newpage
number-sections: true
bibliography: references.bib
---

@@

\newpage
# Introduction
Cycling infrastructure plays a key role in urban transportation, yet quantifying its impact on cycling adoption remains challenging. While previous research demonstrates that protected bike lanes can reduce cycling injuries by up to 89% [@harris2020lane] and increase ridership by up to 171% on specific routes [@nacto2016], less is known about how different types of infrastructure improvements affect bike share system utilization. This knowledge gap is particularly relevant for Toronto, where the city previously committed to significant cycling infrastructure expansion while simultaneously growing its bike share network from 2,854 stations in 2016 to over 6,850 stations in 2020 [@toronto_bike_share_expansion]. Understanding this relationship is essential as cities increasingly rely on bike share systems to complement public transit and provide sustainable transportation options.

Our study estimates the causal effect of cycling infrastructure changes on Bike Share Toronto ridership between 2019-2021, focusing specifically on how different types of infrastructure improvements (protected lanes, on-road lanes, and shared roadways) influence system usage. We examine both new installations and upgrades to existing cycling infrastructure to identify treatment effects through a difference-in-differences design. By combining detailed ridership data with infrastructure locations, we can isolate the impact of specific improvements while controlling for broader trends in cycling adoption.

Our analysis indicates that infrastructure improvements significantly increase bike share ridership, with effects varying substantially by implementation approach. While we observe modest initial increases of 32.4 rides per month during the treatment year, post-implementation effects grow substantially to 133.1 additional monthly rides per bikeway (p < 0.01). In aggregate, this translates to approximately 300,000 additional annual rides across all 186 treated segments - a significant impact given that these improvements represent only 12.6% of Toronto's total bikeways. Protected bike lanes show particularly strong effects, generating 381.9 additional monthly rides post-implementation (p < 0.001) - more than twice the effect of on-road lanes (169.2 rides, p < 0.001) and triple that of shared roadways (111.7 rides, p < 0.001). Notably, these gains occurred despite treated bikeways having higher initial ridership at 209.03 monthly rides, compared to control bikeways 85.59 monthly rides pre-treatment. This suggests infrastructure improvements can substantially increase cycling activity even on popular routes.

These findings have important implications for urban transportation policy as cities work to expand sustainable mobility options. Toronto's commitment to add 100 kilometers of new cycling infrastructure between 2025 and 2027 [@cyclingNetworkPlan] represents a significant opportunity to enhance cycling adoption, particularly if implemented with attention to the differential impacts of various infrastructure types. The substantial ridership gains observed from improving just 12.6% of the network suggest that targeted infrastructure investments can have outsized impacts on system utilization. Furthermore, our finding that improvements to already-popular routes generated significant additional ridership challenges the assumption that infrastructure investments should prioritize underserved areas, highlighting the complex trade-offs cities face when allocating limited resources for cycling infrastructure.

@@

# Data {#sec-data}
## Overview
This study draws on three datasets from the City of Toronto's Open Data Portal [@torontoOpenData]: the Cycling Network (Bikeways) data [@bikeWayData], and both Bike Share Ridership and Stations data [@bikeShareData]. Together, these datasets provide a detailed view of Toronto's cycling landscape, capturing infrastructure development from 2001-2023 and ridership patterns from 2017-2023. While alternative cycling metrics exist, such as manual traffic counts and automated counter data, these sources suffer from inconsistent maintenance and ceased updates after 2019. Our selected datasets offer better temporal coverage and data quality, with regular updates ensuring current, reliable measurements of both infrastructure changes and system usage.

All data cleaning and analysis was conducted using R [@RCoreTeam2023], with key packages including `tidyverse` [@tidyverse], `dplyr` [@dplyr], `readr` [@readr], `jsonlite` [@jsonlite], `arrow` [@arrow], and `sf` [@sf] for spatial analysis.

@@

A vital methodological choice was the use of relative timing rather than calendar years. Due to bikeways having different upgrade or installation years, we align their treatment event year using a relative period. For treated bikeways, this creates a 60-month window: 24 months pre-treatment, 12 months during the treatment year, and 24 months post-treatment. Control bikeways receive randomly assigned pseudo-treatment years (2019, 2020, or 2021) to enable parallel analysis. This approach allows us facilitate direct comparison between treatment and control groups while accounting for the temporal distribution of infrastructure changes.

@@

## Outcome Variables
Our primary outcome variable is seasonally-adjusted monthly ridership for each bikeway (`monthly_rides_adj`). @fig-avg-monthly-ridership shows several important patterns in ridership behavior. The raw ridership (black line) displays pronounced seasonal fluctuations, with consistent peaks during summer months and troughs in winter, reflecting Toronto's strong cycling seasonality. After applying seasonal adjustments (blue line), these cyclical patterns are smoothed while preserving the underlying growth trends. Both treatment and control groups show a general upward trend in ridership over the observation period, suggesting broader growth in cycling activity across Toronto's network.

@@

@fig-infra-type presents distinct infrastructure patterns between groups. While protected lanes dominate both treatment and control groups, comprising approximately 50% of bikeways in each, the distribution of other types differs notably. The treatment group shows a higher proportion of on-road lanes relative to shared roadways, suggesting a deliberate strategy of upgrading less protected infrastructure. This aligns with Toronto's broader policy goal of enhancing cyclist safety through infrastructure improvements [@toronto_council_agenda].

@@

@tbl-summary-statistics in the Appendix provides detailed quantitative breakdowns of both our outcome variables (showing pre/post/treatment period ridership counts and adjusted means) and treatment variables (showing the distribution of protected lanes, on-road lanes, and shared roadways across treatment and control groups).

@@

While Regression Discontinuity Design (RDD) was considered as an alternative approach, our data's temporal granularity (only observing treatment years rather than specific dates) and the gradual nature of infrastructure effects made DiD more appropriate for capturing the full impact of these changes. Moreover, the lack of a clear "forcing variable" that determines treatment status and the potential anticipation effects during construction periods would violate key RDD assumptions, making it unsuitable for our analysis.

Our DiD analysis is ran in R [@RCoreTeam2023] using the `fixest` package [@fixest], which provides estimation methods for fixed effects models. This package is particularly suitable for our analysis as it handles multiple fixed effects efficiently and provides robust standard errors clustered at the bikeway level.

@@

### Parallel Trends Test
An important assumption of the DiD design is that treated and control bikeways would have followed parallel trends in the absence of treatment [@cuborican2024econometrics]. While this counterfactual cannot be directly tested, we can examine pre-treatment trends to assess the plausibility of this assumption. We implement this test by estimating:

@@

### Heterogeneity Analysis
We extend our analysis to examine treatment effect heterogeneity across different types of infrastructure changes. This investigation is necessary for understanding which interventions are most effective at increasing ridership. We estimate two separate models:

@@

These heterogeneity analyses shows important distinctions in the effectiveness of different infrastructure strategies. The particularly strong effect of protected lanes suggests that safety and separation from vehicle traffic may be key factors in encouraging Bike Share usage. Additionally, the stronger effects of upgrades compared to new installations may indicate the importance of improving infrastructure in areas where cycling activity already exists.

# Discussion {#sec-discussion}
Our difference-in-differences analysis of Toronto's cycling infrastructure changes between 2019-2021 demonstrates how different types of cycling infrastructure improvements influence Bike Share utilization. The results demonstrate that while infrastructure investments broadly increase cycling adoption, the magnitude and timing of these effects vary substantially by implementation approach and infrastructure type.

@@

The significant increase in the post-treatment period – to 133.1 additional rides per month (p < 0.01) – shows the true potential of infrastructure improvements once fully integrated into the transportation network. This impact from pre-treatment to post-treatment periods aligns with research from the University of Colorado Denver showing that "building safe facilities for cyclists is one of the biggest factors in road safety for everyone" [@MARSHALL2019100539]. The magnitude of this change suggests that infrastructure improvements don't just make existing cyclists' trips safer, they also alter the decisions behind mode of transportation for many residents.

@@

### Heterogenous Effect Findings
Our heterogeneous effects analysis show which interventions most effectively promote cycling. Protected lanes show remarkably strong effects, generating 381.9 additional monthly rides post-treatment (p < 0.001) – more than twice the impact of on-road lanes (169.2 rides, p < 0.001) and more than triple the effect of shared roadways (111.7 rides, p < 0.001). This substantial difference in effectiveness likely stems from:

@@

The comparison between upgrades and new installations provides equally important discoveries for implementation strategy. Upgrades to existing infrastructure generate more monthly ridership (152.7 adjusted rides), compared to newly-installing bike lanes (115.5 adjusted rides) post-treatment. Additionally, during the treatment period, upgrades to existing infrastructure show immediate effect to ridership (43.8 rides, p < 0.10), whereas new installations show no significant treatment-year effect. This pattern suggests that improving existing cycling routes may be more effective than expanding into new areas, particularly for short-term ridership gains. This finding aligns with Toronto's strategic focus during the same period as our study, on upgrading existing infrastructure to "higher-order" facilities, as outlined in the 2019-2021 Cycling Network Plan Update [@toronto_council_agenda].

@@

- **User surveys**: Qualitative research could help explain the mechanisms behind the stronger effects of protected lanes and how infrastructure influences mode choice decisions. Such research could particularly focus on understanding the "interested but concerned" demographic identified by @nacto2016 as vital for expanding cycling adoption.
- **Network effects**: Future studies could examine how the benefits of infrastructure improvements spill over onto connecting routes and affect system-wide ridership patterns. This could help inform Toronto's goal of creating a more connected cycling network, as outlined in the city's 2025-2027 implementation program [@cyclingNetworkPlan].
- **Neighborhood-level analysis**: Future research could examine how infrastructure effects vary across Toronto's diverse neighborhoods. Such analysis would be particularly beneficial as Toronto expands Bike Share into new areas, helping identify where different types of infrastructure improvements might have the greatest impact. This geographic heterogeneity analysis could also inform equity considerations in infrastructure deployment, ensuring benefits are distributed across the city's diverse communities.
- **Integration with public transit**: Studies could investigate how cycling infrastructure near transit hubs affects multi-modal trip patterns, particularly relevant given Toronto's efforts to integrate Bike Share with the TTC system [@ttcIntegration].

This research contributes to our understanding of how cycling infrastructure influences urban mobility patterns. The clear relationship between infrastructure quality and ridership, particularly the strong effect of protected lanes, provides evidence-based guidance for cities working to encourage sustainable transportation. As Toronto continues expanding its cycling network, these findings can help inform investment priorities and implementation strategies to maximize the impact of infrastructure improvements.

@@

Having this idealized survey would help us directly answer our research question by creating controlled variation in our sampling strategy, comparing treatment areas with new or upgraded bike lanes against control areas without changes. This design allows us to measure both first-stage effects of infrastructure changes on perceptions and intended behavior, and second-stage effects on actual ridership through follow-up surveys, while controlling for demographic and geographic factors. Through carefully designed questions about safety perceptions, route preferences, and time-value tradeoffs, we can identify the specific mechanisms through which infrastructure improvements affect ridership patterns. Additionally, our sampling of different user types enables analysis of heterogeneous effects across current, former, and potential users in various neighborhoods, providing a extensive understanding of how infrastructure changes influence Bike Share adoption.

@@

"Thank you for your time and participation..."Thank you for completing this survey! Your responses will help us better understand how bike infrastructure influences transportation choices in Toronto.

@@

As recommended by Stantcheva [-@Stantcheva2023] for testing persistence of effects, we will implement a three-month follow-up survey. The key objective is to assess whether the stated preferences and intentions from the initial survey translate into actual behavioral changes, while also measuring any infrastructure-induced effects on ridership patterns. To maximize response rates, we will offer an additional $10 gift card incentive and keep the follow-up survey brief (under 3 minutes).

@@

## Tradeoffs and Limitations
Our survey methodology faces several important limitations that could affect the validity of our results. First, self-selection and non-response bias may skew our sample toward respondents who are already interested in cycling or have strong opinions about infrastructure, while our incentive structure might attract particular demographic groups, potentially missing key segments of the population. Our reliance on hypothetical scenarios and stated preferences introduces additional uncertainty, as actual behavior may differ significantly from reported intentions, especially regarding route selection and infrastructure usage. Despite careful design to minimize social desirability bias, respondents may still feel pressure to express support for sustainable transportation initiatives or understate safety concerns. Furthermore, temporal limitations of our three-month follow-up period may inadequately capture seasonal variations in cycling behavior and long-term adaptation to infrastructure changes, while the results may not generalize well to other urban contexts or represent the full diversity of potential users. These limitations suggest that while our survey can provide imporant findings into the relationship between infrastructure and ridership, findings should be interpreted alongside other data sources and methodologies.
